Predicting Post-Election FAANG Stock Prices: Accurate November Trends and Flat December Projections

### 2024-12-03 paper/paper.qmd
committing rendered paper file

---
title: "Predicting Post-Election FAANG Stock Prices: Accurate November Trends and Flat December Projections"
subtitle: "An XGBoost Model Analysis of Market Behavior and Real-World Implications for Meta, Amazon, Apple, Netflix, and Google"
author: 
  - Jimin (Jamie) Lee
thanks: "Code and data are available at: https://github.com/jamiejiminlee/FAANG-Stock-Forecast.git."
date: today
date-format: long
abstract: "This paper presents a predictive model for FAANG stock prices (Meta, Amazon, Apple, Netflix, and Google) during November and December 2024. Using historical stock price data from Yahoo Finance and an XGBoost regression framework, we estimate adjusted stock prices based on features such as lagged prices, moving averages, and volatility. Our findings indicate strong predictive accuracy for short-term trends in November, with predictions closely aligning with actual prices for stocks like AAPL and GOOGL. However, discrepancies, such as the underestimation of NFLX prices, highlight the impact of unique stock-specific dynamics. Predictions for December show a flat trajectory, potentially influenced by data limitations and post-election market sentiment. This analysis provides valuable insights into FAANG stock behavior during critical periods, equipping investors and policymakers to better understand market responses to external events like elections."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| include: false
#| warning: false
#| message: false

@@

- **`Lag_1`**: The closing price from the previous trading day, derived using the `lag()` function. This variable captures sequential patterns in stock prices.
- **`Rolling_Mean_7`**: A 7-day moving average of adjusted closing prices, computed using the `rollmean()` function. This variable smooths short-term fluctuations to highlight weekly trends.
- **`sma_20`**: The 20-day simple moving average (SMA) of adjusted prices, reflecting medium-term price trends.
- **`sma_50`**: The 50-day simple moving average of adjusted prices, providing insights into longer-term trends.
- **`volatility`**: The 20-day rolling standard deviation of adjusted prices, capturing short-term price variability.
- **`daily_return`**: The percentage change in adjusted closing prices relative to the previous day, calculated as:
  $$
  \text{daily_return} = \frac{\text{adjusted}}{\text{Lag_1}} - 1
  $$
  This variable normalizes price changes for comparability across time and stocks.
- **`symbol_encoded`**: A numeric encoding of the stock ticker symbol, allowing the categorical `symbol` variable to be included in the modeling process.

@@

```{r}
# Filter November data for prediction
nov_test <- analysis_data %>%
  filter(as.Date(date) >= as.Date("2024-11-01") & as.Date(date) <= as.Date("2024-11-30"))

@@

```{r}
# Filter December data for prediction
dec_test <- analysis_data %>%
  filter(as.Date(date) >= as.Date("2024-12-01") & as.Date(date) <= as.Date("2024-12-31"))

@@

## US Election Effects
The table illustrates the stock prices of FAANG companies one trading day before and after the elections for the years 2016, 2020, and 2024. The post-election period often shows noticeable fluctuations in stock prices, with some stocks experiencing a sharp increase and others a decrease. For example, AAPL saw a slight increase in 2016 and 2020, while AMZN and NFLX both exhibited declines in 2016. In contrast, the 2024 data reveals an overall upward trend across all FAANG stocks, with particularly strong gains for NFLX, which rose from $755.51 before the election to $780.21 after the election, reflecting a surge in stock prices.

```{r}
library(tidyquant)
library(dplyr)
library(tidyr)
library(knitr)

@@

# Display the table
kable(final_table, caption = "FAANG Stock Prices for One Trading Day Before and After the Election (2016, 2020, and 2024)")

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

---
title: "Predicting Post-Election FAANG Stock Prices: Accurate November Trends and Flat December Projections"
subtitle: "An XGBoost Model Analysis of Market Behavior and Real-World Implications for Meta, Amazon, Apple, Netflix, and Google"
author: 
  - Jimin (Jamie) Lee
thanks: "Code and data are available at: https://github.com/jamiejiminlee/FAANG-Stock-Forecast.git."
date: today
date-format: long
abstract: "This paper presents a predictive model for FAANG stock prices (Meta, Amazon, Apple, Netflix, and Google) during November and December 2024. Using historical stock price data from Yahoo Finance and an XGBoost regression framework, we estimate adjusted stock prices based on features such as lagged prices, moving averages, and volatility. Our findings indicate strong predictive accuracy for short-term trends in November, with predictions closely aligning with actual prices for stocks like AAPL and GOOGL. However, discrepancies, such as the underestimation of NFLX prices, highlight the impact of unique stock-specific dynamics. Predictions for December show a flat trajectory, potentially influenced by data limitations and post-election market sentiment. This analysis provides valuable insights into FAANG stock behavior during critical periods, equipping investors and policymakers to better understand market responses to external events like elections."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

@@

- **`Lag_1`**: The closing price from the previous trading day, derived using the `lag()` function. This variable captures sequential patterns in stock prices.
- **`Rolling_Mean_7`**: A 7-day moving average of adjusted closing prices, computed using the `rollmean()` function. This variable smooths short-term fluctuations to highlight weekly trends.
- **`sma_20`**: The 20-day simple moving average (SMA) of adjusted prices, reflecting medium-term price trends.
- **`sma_50`**: The 50-day simple moving average of adjusted prices, providing insights into longer-term trends.
- **`volatility`**: The 20-day rolling standard deviation of adjusted prices, capturing short-term price variability.
- **`daily_return`**: The percentage change in adjusted closing prices relative to the previous day. This value is calculated by dividing the adjusted closing price by its value from the previous day and subtracting 1. This variable normalizes price changes, making them comparable across time and stocks.
- **`symbol_encoded`**: A numeric encoding of the stock ticker symbol, allowing the categorical `symbol` variable to be included in the modeling process.

@@

# Filter November data for prediction
nov_test <- analysis_data %>%
  filter(as.Date(date) >= as.Date("2024-11-01") & as.Date(date) <= as.Date("2024-11-30"))

@@

# Filter December data for prediction
dec_test <- analysis_data %>%
  filter(as.Date(date) >= as.Date("2024-12-01") & as.Date(date) <= as.Date("2024-12-31"))

@@

## US Election Effects

@tbl-election-effect illustrates the stock prices of FAANG companies one trading day before and after the elections for the years 2016, 2020, and 2024. The post-election period often shows noticeable fluctuations in stock prices, with some stocks experiencing a sharp increase and others a decrease. For example, AAPL saw a slight increase in 2016 and 2020, while AMZN and NFLX both exhibited declines in 2016. In contrast, the 2024 data reveals an overall upward trend across all FAANG stocks, with particularly strong gains for NFLX, which rose from $755.51 before the election to $780.21 after the election, reflecting a surge in stock prices.

```{r}
#| echo: false
#| eval: true
#| label: tbl-election-effect
#| tbl-cap: "FAANG Stock Prices for One Trading Day Before and After the Election (2016, 2020, and 2024)"
#| warning: false

library(tidyquant)
library(dplyr)
library(tidyr)
library(knitr)

@@

# Display the table
kable(final_table)

@@

@@
